---
title: "hw03-Meiqiyu"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The goal is to manipulate and explore a dataset with the dplyr package, complemented by visualizations made with ggplot2. 

```{r}
library(tidyverse)
library(gapminder)
```

# 1.Get the maximum and minimum of GDP per capita for all continents.

The Maximum and minimum of GDP per capita for each continent is listed below.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(gdpPercap_max= max(gdpPercap),
            gdpPercap_min= min(gdpPercap)) 
```

Meanwhile, it's a good idea to show the result in a graph. I list the following two plots both indicating the maximum and minimum value, and the first graph seems better for this question.
```{r}
ggplot(gapminder,aes(continent,gdpPercap))+
  geom_point(aes(color=continent))
```


```{r}
ggplot(gapminder,aes(gdpPercap))+
  geom_histogram(aes(fill=continent))
```

# 2.Look at the spread of GDP per capita within the continents.

The spread of GDP per capita within the continents are listed below.
```{r}
gapminder %>% 
group_by(continent) %>% 
select(continent,gdpPercap)
```

The first graph is a violin plot which shows the spread of GDP per capita within the continents clearly.
```{r}
ggplot(gapminder, aes(continent, gdpPercap, fill=continent)) +  
  geom_violin()+
  geom_jitter(alpha=0.05)
```

The second graph contains five histograms for each continent. 
```{r}
ggplot(gapminder,aes(gdpPercap))+
  facet_grid(~continent)+
  geom_histogram()+
  theme(axis.text.x = element_text(angle = 90))
# use the theme function to rotate x axis to avoid overlapping
```

# 3.How is life expectancy changing over time on different continents?
In the table below, the data is organized with year growing for each continent, so the trends of lifeExp on different continents can be seen clearly.
```{r}
gapminder %>% 
  group_by(continent) %>% 
  select(continent,year,lifeExp)
```

The graph cotains five point plots showing the lifeExp over year on different continents.

```{r}
ggplot(gapminder,aes(year,lifeExp))+
  facet_wrap(~continent, scales = "free_x")+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90))
# use the theme function to rotate x axis to avoid overlapping
```



                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          











